---
title: "Getting Started with hsir"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting Started with hsir}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
--- 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  echo = TRUE
)
```

# Getting Started with hsir

The **hsir** package provides lightweight, fast and memory-efficient functions for image recognition and processing tasks. This vignette will guide you through the main features and usage patterns.

## Installation

```r
# Install from GitHub
devtools::install_github("JannikVindeloev/hsir")

# Load the package
library(hsir)
```

## Package Overview

The hsir package is designed with the following principles:

1. **Memory Efficiency**: Uses optimized data structures like `data.table`
2. **Speed**: Leverages fast matrix operations from `matrixStats` and `fastmatrix`
3. **Tidyverse Compatibility**: Follows tidyverse conventions for consistent API
4. **Comprehensive**: Includes utilities for image data manipulation and processing

## Core Functions

### Package Availability Check

Before using the package, you can check if all required dependencies are available:

```{r, check-packages}
# Check if all required packages are available
check_packages()
```

### Image Data Structures

The `create_image_dt()` function creates an optimized data structure for storing image data:

```{r, image-dt}
# Create sample image data
img_data <- matrix(1:100, ncol = 10)

# Convert to optimized data.table
img_dt <- create_image_dt(img_data)

# View the structure
str(img_dt)
head(img_dt)
```

For multi-dimensional arrays (e.g., RGB images):

```{r, image-dt-array}
# Create 3D array (e.g., RGB image)
rgb_data <- array(1:24, dim = c(2, 3, 4))

# Convert to optimized data.table
rgb_dt <- create_image_dt(rgb_data)

# View the structure
str(rgb_dt)
head(rgb_dt)
```

### Fast Matrix Operations

The `fast_matrix_ops()` function performs optimized matrix operations for image processing:

```{r, matrix-ops}
# Create sample matrix
mat <- matrix(1:100, ncol = 10)

# Calculate row means
row_means <- fast_matrix_ops(mat, "mean")
row_means

# Calculate column medians
col_medians <- fast_matrix_ops(t(mat), "median")
col_medians

# All available operations
operations <- c("mean", "median", "sum", "min", "max")
results <- lapply(operations, function(op) fast_matrix_ops(mat, op))
names(results) <- operations
results
```

### Handling NA Values

The functions are designed to handle NA values efficiently:

```{r, na-handling}
# Create matrix with NA values
mat_na <- matrix(c(1:9, NA), ncol = 2)
mat_na

# Calculate means ignoring NA values
fast_matrix_ops(mat_na, "mean", na.rm = TRUE)

# Calculate means including NA values
fast_matrix_ops(mat_na, "mean", na.rm = FALSE)
```

## Performance Considerations

The hsir package is optimized for performance:

1. **Memory Usage**: Uses data.table which is more memory efficient than data.frame
2. **Speed**: Leverages compiled C/C++ code from matrixStats and fastmatrix
3. **Vectorization**: Uses vectorized operations where possible
4. **Minimal Copying**: Avoids unnecessary data copying

## Best Practices

1. **Use data.table**: For large datasets, always use data.table for memory efficiency
2. **Matrix Operations**: Use the provided matrix operations instead of base R functions
3. **Avoid Loops**: Use vectorized operations or apply functions instead of loops
4. **Memory Management**: Be mindful of memory usage with large images

## Example Workflow

Here's a complete example workflow for processing image data:

```{r, workflow}
# Load required packages
library(hsir)
library(data.table)

# Simulate image data (100x100 pixels)
set.seed(123)
image_data <- matrix(rnorm(10000), ncol = 100)

# Convert to optimized data structure
image_dt <- create_image_dt(image_data)

# Calculate statistics for each row (pixel row)
row_stats <- data.table(
  mean = fast_matrix_ops(image_data, "mean"),
  median = fast_matrix_ops(image_data, "median"),
  min = fast_matrix_ops(image_data, "min"),
  max = fast_matrix_ops(image_data, "max"),
  sum = fast_matrix_ops(image_data, "sum")
)

# View the results
head(row_stats)

# Calculate overall statistics
overall_stats <- list(
  mean = mean(image_data, na.rm = TRUE),
  sd = sd(image_data, na.rm = TRUE),
  min = min(image_data, na.rm = TRUE),
  max = max(image_data, na.rm = TRUE)
)

overall_stats
```

## Conclusion

The hsir package provides a comprehensive set of tools for efficient image processing in R. By leveraging optimized data structures and algorithms, it offers significant performance improvements over traditional approaches.

For more information, see the package documentation (`?hsir`) and the README file.
